---
bibliography: references.bib
csl: "apa-6th-edition.csl"
output: 
   html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Climate Security

*Climate security* refers to any **threats or risks to states, societies, and individual citizens directly or indirectly caused by climate change**. Climate variability and climate extreme events are impacting food, land and water systems thereby threatening agriculture productivity, and diminishing natural resources of land and water, which increases resource-based conflicts. 


### Ethiopia

Located in the horn of Africa.

```{r E1, message=F, warning=F, echo=F}
library(terra)
root <- 'D:/OneDrive - CGIAR/Analysis/CSO/'
iso <- 'ETH'
conf <- geojsonio::geojson_sf(paste0(root,iso,'/clim_conflict_ips_overlays.geojson'))
bdy1 <- geodata::gadm(country=iso, level=1, path=paste0(root,iso,'/'))
bdy1 <- sf::st_as_sf(bdy1)
bdy2 <- geodata::gadm(country=iso, level=2, path=paste0(root,iso,'/'))
bdy2 <- sf::st_as_sf(bdy2)
bdy3 <- geodata::gadm(country=iso, level=3, path=paste0(root,iso,'/'))
bdy3 <- sf::st_as_sf(bdy3)
library(tmap)
library(mapview)
tmap_mode("view")
et <- tm_shape(bdy1, name = 'Region')+
  tm_borders() +
  tm_shape(bdy2, name = 'Zone')+
  tm_borders() +
  tm_shape(bdy3, name = 'Woreda')+
  tm_borders() +
  tm_shape(bdy3, name = 'Woreda')+
  tm_format("World")
et 
```



## References

