---
title: "Data Science"
author: "Benson Kenduiywo"
date: '2022-08-11'
csl: "apa-6th-edition.csl"
output: 
   html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: false
editor_options: 
  chunk_output_type: console
---

```{r, global_options, tidy = TRUE, echo=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60))
opts_knit$set(global.par = TRUE)
```

## Background

This tutorial seeks to quantify how climate can influence the likelihood and intensity of conflict via its impact on socio-economic vulnerabilities, such as food and nutrition insecurity, forced displacement and migration, inequality, etc. The Spatial Analysis is an important component of this approach providing answers to the following research questions:

-	Where are the most vulnerable areas to climate induced insecurities and risks?  
-	Who are vulnerable groups to climate and security risks that should be targeted? 

## Introduction

This exercise explores the how of different climate influences the intensity different types of conflicts. Regional data from [ACLED](https://acleddata.com/curated-data-files/#aggregated) is adopted. ACLED is an event-based data project designed for disaggregated conflict analysis and crisis mapping.

Prepare environment and Load the data

```{r d1}
rm(list=ls(all=TRUE))
root <- "/home/servir/RHEAS/data/conflict/"
temp <- read.csv(paste0(root, "Africa_1997-2022_Aug05.csv"), stringsAsFactors =  FALSE)

knitr::kable(head(temp, n=5), booktabs=TRUE) 
```

Filter conflict events for Kenya only

```{r d2}
con <- subset(temp, COUNTRY=="Kenya")
```

## Visualization

### Conflict types

Next lets visualize conflicts type and fatalities.

```{r v1}
cause <- tapply(con$FATALITIES, con$EVENT_TYPE, sum, na.rm=TRUE)
x <- sort(round(cause))
par(mai=c(0.5,3,0.5,1)) #c(bottom, left, top, right)
barplot(x, horiz=T, las=1)
```

### Spatial distribution

Violence against civilians is the leading type of conflict. Let's visualize this event spatially. To do that the dataframe should be converted to a `SpatialPointsDataFrame`.

```{r v2}
df1 <- con
library(sp)
coordinates(df1) <- ~LONGITUDE+LATITUDE
proj4string(df1) <- CRS("+proj=longlat +datum=WGS84")

```

Visualize the spatial object.

```{r v3, echo=F , message=F}
library(tmap)
library(mapview)
tmap_mode("view")
map_a <- tm_shape(df1, name="Violence against civilians") +
  tm_bubbles(size="FATALITIES", col = "EVENT_TYPE") +
  tm_format("World")
map_a

lf <- tmap_leaflet(map_a)
mapshot(lf, file = "CivilianViolencemap.png")

```

Summary of conflicts distribution:

1) Battles are dominant in then NW, 
2) Many explosions are reported in the Nairobi county.
3) Riots are dominant in the Nairobi, Kisumu, Machakos and Mombasa.
4) The other forms seem wide spread.

Let's explore more violence against civilians.

```{r v4, echo=F , message=F}
df1 <- subset(df1, EVENT_TYPE=="Violence against civilians")
tmap_mode("view")
map_b <- tm_shape(df1, name="Violence against civilians") +
  tm_bubbles(size="FATALITIES", col = "SUB_EVENT_TYPE") +
  tm_format("World")
map_b
lf <- tmap_leaflet(map_a)
mapshot(lf, file = "CivilianViolencemap_b.png")

#Save for later use
saveRDS(df1, "civilianViolence.rds")
```

Violence against civilians in the form of attack are dominant and widespread across the country. Many cases are reported in Turkana and towards the Eastern side of the country.

### Temporal distribution

When do these form of violence happen?

```{r v5}
library(reshape2)
temp <- subset(df1@data, select = c(YEAR, FATALITIES, SUB_EVENT_TYPE))
temp <- aggregate(FATALITIES~SUB_EVENT_TYPE+YEAR, temp, sum, na.rm=T)
temp <- reshape(temp, direction="wide", idvar="SUB_EVENT_TYPE", timevar="YEAR")
colnames(temp) <- gsub("FATALITIES.", "", colnames(temp))
temp <- as.matrix(temp)
rownames(temp) <- temp[,1]
cola <- c("red", "green", "blue")
par(mfrow=c(1, 1), mar=c(5, 5, 4, 5))
barplot(temp[,-1], legend =  rownames(temp), las=2, args.leg=list(x = "topright", bty = "n", inset=c(-0.2, 0), cex=0.75), col=cola, ylab="Fatalities")

```

From the figure the years 1997, 1998, 2000, 2007, 2008, 2013, and, 2017 had the highest attacks with Kenya civilians being the most vulnerable. 

In the next section we integrate food production indices to relate evaluate any relationship between food production and conflicts against civilians.

[**NEXT PAGE**](modis.html).



