---
title: "Data Science"
author: "Benson Kenduiywo"
date: '2022-08-11'
csl: "apa-6th-edition.csl"
output: 
   html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: false
editor_options: 
  chunk_output_type: console
---

```{r, global_options, tidy = TRUE, echo=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60))
opts_knit$set(global.par = TRUE)
```

## Introduction

This exercise explores the how of different climate influences the intensity different types of conflicts. Regional data from [ACLED](https://acleddata.com/curated-data-files/#aggregated) is adopted. ACLED is an event-based data project designed for disaggregated conflict analysis and crisis mapping.


Prepare environment and Load the data

```{r d1}
rm(list=ls(all=TRUE))
root <- "/home/servir/RHEAS/data/conflict/"
temp <- read.csv(paste0(root, "Africa_1997-2022_Aug05.csv"), stringsAsFactors =  FALSE)

knitr::kable(head(temp, n=5), booktabs=TRUE) 
```

Filter conflict events for Kenya only

```{r d2}
con <- subset(temp, COUNTRY=="Kenya")
```

## Visualization

Next lets visualize conflicts type and fatalities.

```{r v1}
cause <- tapply(con$FATALITIES, con$EVENT_TYPE, sum, na.rm=TRUE)
x <- sort(round(cause))
par(mai=c(0.5,3,0.5,1)) #c(bottom, left, top, right)
barplot(x, horiz=T, las=1)
```

Violence against civilians is the leading type of conflict. Let's visualize this event spatially.



```{r}
if (!"raster" %in% installed.packages()){
  install.packages("raster", dependencies = T)
}
library(raster)
ken <- getData("GADM", country="KEN", level=1)
plot(ken)
text(ken, toupper(ken$NAME_1),  cex=0.6)
```




