---
bibliography: references.bib
csl: "apa-6th-edition.csl"
output: 
   html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: false
editor_options: 
  chunk_output_type: console
---

```{r, global_options, tidy = TRUE, echo=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60))
opts_knit$set(global.par = TRUE)
```

## Climate metrics

Load necessary packages and define directories.

```{r refdata, echo=FALSE, warning=FALSE, message=FALSE}
rm(list = ls(all=TRUE))
unlink(".RData")
library(dplyr)
library(reshape2)
root <- "/home/servir/RHEAS/data/conflict/"
metric <- paste0(root, "metric")
ref <- paste0(root, "ref/")

```


### Spatial metrics

Load Gross Primary Productivity (GPP) MODIS (2000--2022) and ERA5 (1997--2020) climate metrics computed in [Earth Engine]([here](https://code.earthengine.google.com/29f887f9a190abdb7aef6b8bab5b3ccf)) and convert to long format. These metrics have been spatially aggregated to county boundaries as depicted by Figure 4 in [@Kenduiywo_ISPRS_2020].

```{r modis, echo=FALSE, warning=FALSE, message=FALSE}
wide2long <- function(df, variable){
  df <- subset(df, select =-.geo)
  df <- subset(df, NAME_1 != "Counties")
  melt(df[, -1], variable.name = "date", 
          value.name = variable, id.vars = "NAME_1")
}

#1.0 GPP
temp <- list.files(metric, pattern=glob2rx("*gpp*"), full.names = T)
gpp <-  read.csv(temp, stringsAsFactors =  FALSE)
gpp <-  wide2long(gpp, "gpp")

#2.0 air_temp
temp <- list.files(metric, pattern=glob2rx("*temp*"), full.names = T)
tmp <-  read.csv(temp, stringsAsFactors =  FALSE)

tmp <-  wide2long(tmp, "temp")

#3.0 total daily precipitation
temp <- list.files(metric, pattern=glob2rx("*precip*"), full.names = T)
prec <-  read.csv(temp, stringsAsFactors =  FALSE)

prec <-  wide2long(prec, "precip")

```

Format the date variable by removing unnecessary characters.

```{r y1, echo=FALSE, warning=FALSE, message=FALSE}
formatDate <- function(m, weather=F){
  if(weather){
    m$date <- gsub("X","", m$date)
    #m$date <- gsub("_nd","", m$date)
    m$date <- substr(m$date,1,8)
    m$date <- as.Date(as.character(m$date), format = "%Y%m%d")
    m[1,]
    colnames(m)[1] <- "County"
    
  }else{
    m$date <- gsub("X","", m$date)
    #m$date <- gsub("_nd","", m$date)
    m$date <- substr(m$date,1,10)
    m$date <- as.Date(as.character(m$date), format = "%Y_%m_%d")
    m[1,]
    colnames(m)[1] <- "County"
  }
  
  return(m)
}

gpp <- formatDate(gpp)
knitr::kable(head(gpp, n=2))
tmp <- formatDate(tmp, weather = T)
knitr::kable(head(tmp, n=2))
prec <- formatDate(prec, weather = T)
knitr::kable(head(prec, n=2))
```


## Spatial-temporal metrics

Temporally aggregate the spatial metrics per year to obtain *spatial-temporal climate metrics*. Event of conflicts from violence against civilians will also be aggregated per year and county.

```{r st1}
gpp$date <- format(as.Date(gpp$date, format="%d/%m/%Y"),"%Y")
gpp <- aggregate(.~County+date, data=gpp, mean, na.rm=T)
tmp$date <- format(as.Date(tmp$date, format="%d/%m/%Y"),"%Y")
tmp <- aggregate(.~County+date, data=tmp, mean, na.rm=T)
prec$date <- format(as.Date(prec$date, format="%d/%m/%Y"),"%Y")
prec <- aggregate(.~County+date, data=prec, mean, na.rm=T)

con <- readRDS("civilianViolence.rds")
con <- subset(con@data, select= c(ADMIN1, FATALITIES, YEAR))
names(con)[1] <- "County"
names(con)[3] <- "date"
con <- aggregate(.~County+date, data=con, sum, na.rm=T)

```


Let's make some plots and evaluate trends. First merge each jmetric with conflict data. To do so check for county naming consistencies. 


```{r st2}
con$County <- toupper(con$County)
gpp$County <- toupper(gpp$County)
tmp$County <- toupper(tmp$County)
prec$County <- toupper(prec$County)
c <- sort(unique(con$County))
c[!c %in% sort(unique(gpp$County))]
c[!c %in% sort(unique(tmp$County))]
c[!c %in% sort(unique(prec$County))]
```

Edit the inconsistencies on the identified counties.

```{r st3}
con$County[con$County == "MURANGA"] <- "MURANG'A"
con$County[con$County == "ELGEYO MARAKWET" ] <- "ELGEYO-MARAKWET"
```

We can also compute anomalies for these climate metrics. The _Z-score_ is a good measure to capture anomalies. Z-scores indicate the deviations of seasonal/annual metrics from its long-term annual mean. The unit is “standard deviations”. This a z-score of -1 means that the value is 1 standard deviation below the (expected) mean value i.e.;

```{r st4}
zscore <- function(y){
  (y - mean(y, na.rm=TRUE) ) / (sd(y, na.rm=TRUE))
}

```

Compute the zScored metrics per year per Country

```{r d7}
gpp <- ungroup(mutate(group_by(gpp,County), zgpp=zscore(gpp)))
tmp <- ungroup(mutate(group_by(tmp,County), ztemp=zscore(temp)))
prec <- ungroup(mutate(group_by(prec,County), zprecip=zscore(precip)))
con <- ungroup(mutate(group_by(con,County), zFATALITIES=zscore(FATALITIES)))
```

## Visualize trends

### Simple plots

Make trend plots.

```{r tr1}
x <- aggregate(FATALITIES~date, data=con, mean, na.rm=T)
y <- aggregate(temp~date, data=tmp, mean, na.rm=T)
z <- aggregate(precip~date, data=prec, mean, na.rm=T)
k <- aggregate(gpp~date, data=gpp, mean, na.rm=T)

par(mfrow = c(2, 2), mar = c(4, 5, 1, 1)) #c(bottom, left, top, right)
plot(FATALITIES~date, data=x, type="l", col="red", ylab="# Fatalities", xlab= "Year", cex.axis =1.2, cex.lab = 1.2)
plot(gpp~date, data=k, type="l", col="red", ylab="GPP", xlab= "Year", cex.axis =1.2, cex.lab = 1.2)
plot(precip~date, data=z, type="l", col="red", ylab="Precipitation", xlab= "Year", cex.axis =1.2, cex.lab = 1.2)
plot(temp~date, data=y, type="l", col="red", ylab="2m Air Tempetature", xlab= "Year", cex.axis =1.2, cex.lab = 1.2)

```


### Regression plots

As we proceed let's merge all the datasets to one `data.frame`

```{r merg1}
df_list <- list(con, tmp, gpp, prec)
dff <- Reduce(function(x, y) merge(x, y, by=c("County","date")), df_list)

```

Make regression plots to identify relationship between conflicts and various climate EO metrics.

```{r tr2}
library(ggplot2)
library(gridExtra)
aa <- melt(subset(dff, select=-c(temp, gpp, precip, FATALITIES)), variable.name = "metric_name", value.name = "metric", id.vars = c("County", "date", "zFATALITIES"))

ggplot(aa, aes(metric, zFATALITIES)) +
  geom_point() +
  geom_smooth(method="loess") +
  facet_wrap(~ metric_name)


#Explore other plots with no anomalies
p1 <- ggplot(dff, aes(gpp, FATALITIES)) +
  geom_point() +
  geom_smooth(method = "loess")

p2 <- ggplot(dff, aes(temp, FATALITIES)) +
  geom_point() +
  geom_smooth(method = "loess")

p3 <- ggplot(dff, aes(precip, FATALITIES)) +
  geom_point() +
  geom_smooth(method = "loess")
grid.arrange(p1, p2, p3, ncol = 2)

```


Seems like precipitation is a good indicator of conflicts. For instance, conflicts are high when precipitation is low..


## References


