---
bibliography: references.bib
csl: "apa-6th-edition.csl"
output: 
   html_document:
    toc: true
    toc_float: true
    toc_collapsed: true
    toc_depth: 3
    number_sections: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Climate Earth Observation variables

### MODIS Gross Primary Productivity

Lets use MODIS Gross Primary Productivity (GPP) and evaluate the relationship between violence against civilians and food productivity. GPP is a product from MODIS that was acquired from MOD17 data series products generated from Terra satellite. It is based on the radiation-use efficiency concept and can potentially be used to quantify generation of new biomass in vegetation and a proxy to food production [@Kenduiywo_ISPRS_2020].


GPP 8-day composite is available at 500 m spatial resolution period with product names **MOD17A2H** and **MYD17A2H** in Terra and Aqua respectively [@12_NASA_2021]. Let's download the product which available from 2000. 

```{r m1}
if (!"luna" %in% installed.packages()){
  remotes::install_github("rspatial/luna")
}
library(luna)
prod <- getProducts("^MOD17A2|^MYD17A2")
knitr::kable(table(head(prod, n=3)), align = 'l')

```

We are almost ready to download the data. Let's get Kenya's country boundary and start data downloads. 

```{r m2, message=F, warning=F}
if (!"raster" %in% installed.packages()){
  install.packages("raster", dependencies = T)
}
library(raster)
aoi <- getData("GADM", country="KEN", level=0)
plot(aoi)

```

Define data parameters: product name, start and end date, and area of interest. We define date as follows: 

```{r m3}
start <- "2001-01-01"
end <- "2001-02-12"

```

Thereafter, extract existing data files, this Terra GPP products.

```{r m4}
product <- "MOD17A2H"
mf <- luna::getModis(product, start, end, aoi=aoi, download = FALSE)
length(mf)
```

MODIS data can be download from [EarthData](https://urs.earthdata.nasa.gov/users). This requires specification of a storage directory, a user's name and password. Here, the password is saved in a directory that we will read. For the sake of time it takes to donwload the data a demo is only done for 5 days data. Later Earth Engine will be used to process the data and code link shared for purposes of time constraints.

```{r m5}
pass <- readRDS("earthdata.rds")
modis_path <- "/home/bkenduiywo/R/raw/" 
dir.create(modis_path)
mstart <- "2001-01-01"
mend <- "2001-01-05"
luna::getModis(product, start, end, aoi=aoi, download=TRUE,
                     path=modis_path, username=pass$username,
                     password=pass$password)

```

Further, processing steps are available [here](https://bensonkenduiywo.github.io/dip/MODISprocessing.html). The steps are covered on Earth Engine to save on time those steps were done using as shown [here](https://code.earthengine.google.com/29f887f9a190abdb7aef6b8bab5b3ccf).

[**NEXT PAGE**](metrics.html).

## References

